 extracting environmenal and oceanographic variables

### 2023-05-11 03_environ_var.Rmd
renamed to environ_var, removed code with plots

---
title: "environmenal/exploratory analysis"
output: html_document
date: "2022-11-06"
---

@@

##FIX THIS - try testing on ONE buoy
for (t in tracklist) {
  Hdata <- matchEnvData(as.data.frame(get(paste0("tracks_SS_", substr(t, 11, 12)))), nc=PAMmisc::hycomList, var=c('water_temp', 'salinity'), raw=TRUE, depth = 0:1000)

  thermData <- bind_rows(lapply(Hdata, function(x) {
    # do thermo calcs
    result <- calcThermoVars(x$matchDepth, x$water_temp)
    # get specific values
    result$temp400 <- x$water_temp[x$matchDepth == 400]
    result$sal400 <- x$salinity[x$matchDepth == 400]
    result
}))
  bindr <- cbind(get(paste0("tracks_SS_", substr(t, 11, 12))), thermData)
  assign(paste0("tracks_SS_", substr(t, 11, 12)), bindr) #fix this part?
}
```

```{r}
test08 <- matchEnvData(as.data.frame(tracks_SS_08), nc=PAMmisc::hycomList, var=c('water_temp', 'salinity'), raw=TRUE, depth = 0:1000)

@@

```{r}
#hycomData0 <- matchEnvData(as.data.frame(tracks_SS_08[1:250,]), nc=PAMmisc::hycomList, var=c('water_temp', 'salinity'), raw=TRUE, depth = 0:1000, fileName = "data-raw/hycomm0.nc")

#test <- matchEnvData(as.data.frame(tracks_SS_07), nc=PAMmisc::hycomList, var=c('water_temp', 'salinity'), raw=TRUE, depth = 0:1000)

#hycomData1 <- matchEnvData(as.data.frame(tracks_SS_08[251:500,]), nc=PAMmisc::hycomList, var=c('water_temp', 'salinity'), raw=TRUE, depth = 0:1000, fileName = "data-raw/hycomm1.nc")

#hycomData2 <- matchEnvData(as.data.frame(tracks_SS_08[501:750,]), nc=PAMmisc::hycomList, var=c('water_temp', 'salinity'), raw=TRUE, depth = 0:1000, fileName = "data-raw/hycomm2.nc")

#hycomData3 <- matchEnvData(as.data.frame(tracks_SS_08[751:1200,]), nc=PAMmisc::hycomList, var=c('water_temp', 'salinity'), raw=TRUE, depth = 0:1000, fileName = "data-raw/hycomm3.nc")

#hycomData4 <- matchEnvData(as.data.frame(tracks_SS_08[1201:1500,]), nc=PAMmisc::hycomList, var=c('water_temp', 'salinity'), raw=TRUE, depth = 0:1000, fileName = "data-raw/hycomm4.nc")

#hycomData5 <- matchEnvData(as.data.frame(tracks_SS_08[1501:1800,]), nc=PAMmisc::hycomList, var=c('water_temp', 'salinity'), raw=TRUE, depth = 0:1000, fileName = "data-raw/hycomm5.nc")

#hycomData6 <- matchEnvData(as.data.frame(tracks_SS_08[1801:2145,]), nc=PAMmisc::hycomList, var=c('water_temp', 'salinity'), raw=TRUE, depth = 0:1000, fileName = "data-raw/hycomm6.nc")

#hycomData7 <- matchEnvData(as.data.frame(tracks_SS_08[2146:2469,]), nc=PAMmisc::hycomList, var=c('water_temp', 'salinity'), raw=TRUE, depth = 0:1000, fileName = "data-raw/hycomm7.nc")

#hycomData8 <- matchEnvData(as.data.frame(tracks_SS_08[2470:2769,]), nc=PAMmisc::hycomList, var=c('water_temp', 'salinity'), raw=TRUE, depth = 0:1000, fileName = "data-raw/hycomm8.nc")

# Run the 9 nc lists that make up tracks_SS_08
#hycomData0 <- matchEnvData(as.data.frame(tracks_SS_08[1:250,]), var=c('water_temp', 'salinity'), raw=TRUE, depth = 0:1000, nc = "data-raw/hycomm0.nc")
#hycomData1 <- matchEnvData(as.data.frame(tracks_SS_08[251:500,]), var=c('water_temp', 'salinity'), raw=TRUE, depth = 0:1000, nc = "data-raw/hycomm1.nc")
#hycomData2 <- matchEnvData(as.data.frame(tracks_SS_08[501:750,]), var=c('water_temp', 'salinity'), raw=TRUE, depth = 0:1000, nc = "data-raw/hycomm2.nc")
#hycomData3 <- matchEnvData(as.data.frame(tracks_SS_08[751:1200,]), var=c('water_temp', 'salinity'), raw=TRUE, depth = 0:1000, nc = "data-raw/hycomm3.nc")
#hycomData4 <- matchEnvData(as.data.frame(tracks_SS_08[1201:1500,]), var=c('water_temp', 'salinity'), raw=TRUE, depth = 0:1000, nc = "data-raw/hycomm4.nc")
#hycomData5 <- matchEnvData(as.data.frame(tracks_SS_08[1501:1800,]), var=c('water_temp', 'salinity'), raw=TRUE, depth = 0:1000, nc = "data-raw/hycomm5.nc")
#hycomData6 <- matchEnvData(as.data.frame(tracks_SS_08[1801:2145,]), var=c('water_temp', 'salinity'), raw=TRUE, depth = 0:1000, nc = "data-raw/hycomm6.nc")
#hycomData7 <- matchEnvData(as.data.frame(tracks_SS_08[2146:2469,]), var=c('water_temp', 'salinity'), raw=TRUE, depth = 0:1000, nc = "data-raw/hycomm7.nc")
#hycomData8 <- matchEnvData(as.data.frame(tracks_SS_08[2470:2769,]), var=c('water_temp', 'salinity'), raw=TRUE, depth = 0:1000, nc = "data-raw/hycomm8.nc")
```

Use above function on the smaller hycom datasets (0-6)
```{r}
#Optional plot:
calcThermoVars(test[[1]]$matchDepth, test[[1]]$water_temp, plot=TRUE)

thermData0 <- bind_rows(lapply(hycomData0, function(x) {
    # do thermo calcs
    result <- calcThermoVars(x$matchDepth, x$water_temp)
    # get specific values
    result$temp400 <- x$water_temp[x$matchDepth == 400]
    result$sal400 <- x$salinity[x$matchDepth == 400]
    result
}))

thermData1 <- bind_rows(lapply(hycomData1, function(x) {
    # do thermo calcs
    result <- calcThermoVars(x$matchDepth, x$water_temp)
    # get specific values
    result$temp400 <- x$water_temp[x$matchDepth == 400]
    result$sal400 <- x$salinity[x$matchDepth == 400]
    result
}))

thermData2 <- bind_rows(lapply(hycomData2, function(x) {
    # do thermo calcs
    result <- calcThermoVars(x$matchDepth, x$water_temp)
    # get specific values
    result$temp400 <- x$water_temp[x$matchDepth == 400]
    result$sal400 <- x$salinity[x$matchDepth == 400]
    result
}))

thermData3 <- bind_rows(lapply(hycomData3, function(x) {
    # do thermo calcs
    result <- calcThermoVars(x$matchDepth, x$water_temp)
    # get specific values
    result$temp400 <- x$water_temp[x$matchDepth == 400]
    result$sal400 <- x$salinity[x$matchDepth == 400]
    result
}))

thermData4 <- bind_rows(lapply(hycomData4, function(x) {
    # do thermo calcs
    result <- calcThermoVars(x$matchDepth, x$water_temp)
    # get specific values
    result$temp400 <- x$water_temp[x$matchDepth == 400]
    result$sal400 <- x$salinity[x$matchDepth == 400]
    result
}))

thermData5 <- bind_rows(lapply(hycomData5, function(x) {
    # do thermo calcs
    result <- calcThermoVars(x$matchDepth, x$water_temp)
    # get specific values
    result$temp400 <- x$water_temp[x$matchDepth == 400]
    result$sal400 <- x$salinity[x$matchDepth == 400]
    result
}))

thermData6 <- bind_rows(lapply(hycomData6, function(x) {
    # do thermo calcs
    result <- calcThermoVars(x$matchDepth, x$water_temp)
    # get specific values
    result$temp400 <- x$water_temp[x$matchDepth == 400]
    result$sal400 <- x$salinity[x$matchDepth == 400]
    result
}))

thermData7 <- bind_rows(lapply(hycomData7, function(x) {
    # do thermo calcs
    result <- calcThermoVars(x$matchDepth, x$water_temp)
    # get specific values
    result$temp400 <- x$water_temp[x$matchDepth == 400]
    result$sal400 <- x$salinity[x$matchDepth == 400]
    result
}))

thermData8 <- bind_rows(lapply(hycomData8, function(x) {
    # do thermo calcs
    result <- calcThermoVars(x$matchDepth, x$water_temp)
    # get specific values
    result$temp400 <- x$water_temp[x$matchDepth == 400]
    result$sal400 <- x$salinity[x$matchDepth == 400]
    result
}))

# Bind all thermocline datasets together
thermData <- rbind(thermData0, thermData1, thermData2, thermData3, thermData4, 
                   thermData5, thermData6, thermData7, thermData8)

# Bind main thermocline dataset to main track/soundscape dataframe
tracks_SS_08 <- cbind(tracks_SS_08, thermData)

# New tracks_SS_08 variable contains Hycom variables now
##### FINAL tracks_ss_08 dataframe ######
View(tracks_SS_08)
```
```{r}
thermData <- bind_rows(lapply(test, function(x) {
    # do thermo calcs
    result <- calcThermoVars(x$matchDepth, x$water_temp)
    # get specific values
    result$temp400 <- x$water_temp[x$matchDepth == 400]
    result$sal400 <- x$salinity[x$matchDepth == 400]
    result
}))

test_fin <- cbind(tracks_SS_07, thermData)
print(data)
```
testing hycom bind on ONE drift first
```{r}
test22 <- matchEnvData(as.data.frame(tracks_SS_22), nc=PAMmisc::hycomList, var=c('water_temp', 'salinity'), raw=TRUE, depth = 0:1000)

thermData22 <- bind_rows(lapply(test22, function(x) {
    # do thermo calcs
    result <- calcThermoVars(x$matchDepth, x$water_temp)
    # get specific values
    result$temp400 <- x$water_temp[x$matchDepth == 400]
    result$sal400 <- x$salinity[x$matchDepth == 400]
    result
}))

test_fin <- cbind(tracks_SS_22, thermData22)
View(test_fin)
```

```{r}
saveRDS(tracks_SS_10, 'data/fulltrack10.rda')
saveRDS(tracks_SS_12, 'data/fulltrack12.rda')
saveRDS(tracks_SS_13, 'data/fulltrack13.rda')
saveRDS(tracks_SS_14, 'data/fulltrack14.rda')
saveRDS(tracks_SS_16, 'data/fulltrack16.rda')
saveRDS(tracks_SS_18, 'data/fulltrack18.rda')
saveRDS(tracks_SS_19, 'data/fulltrack19.rda')
saveRDS(tracks_SS_20, 'data/fulltrack20.rda')
saveRDS(tracks_SS_21, 'data/fulltrack21.rda')
saveRDS(tracks_SS_22, 'data/fulltrack22.rda')
saveRDS(tracks_SS_23, 'data/fulltrack23.rda')
saveRDS(tracks_SS_07, 'data/fulltrack07.rda')
saveRDS(tracks_SS_08, 'data/fulltrack08.rda')
```

Get avg daily environmental variables
```{r}
means_08 <- tracks_SS_08 %>%
  group_by(as.Date(UTC))%>%
  summarize(sst_mean = mean(sst_mean),
            dist_slope = mean(dist2slope),
            depth = mean(depth),
            slope = mean(slope),
            dist_shore = mean(dist2shore),
            chlorophyll_mean = mean(chlorophyll_mean),
            ssh_mean = mean(ssh_mean),
            mldDepth = mean(mldDepth),
            mldTemp = mean(mldTemp),
            ttDepth = mean(ttDepth),
            ttTemp = mean(ttTemp),
            temp400 = mean(temp400),
            sal400 = mean(sal400),
            BWpresence = sum(BWpresence))
```

### Exploratory Analysis ######

### Third octave levels histograms

```{r}
# Drift 8 - all TOL bands, facet wrap
pivot <- tracks_SS_08 %>%
  pivot_longer(cols = starts_with("TOL") | starts_with("BB"), names_to = "SL", values_to = "Hz")    # Pivot longer for histogram facet grid

ggplot(data = pivot, aes(Hz, fill = factor(BWpresence))) +
  geom_histogram(binwidth = 3, color = "black") +
  theme_classic() +
  ggtitle("Distribution of third octave level intensities: Drift 8") +
  labs(x = "Sound Intensity (dB)") +
  guides(fill = guide_legend(title = "Beaked Whale Presence")) +
  scale_fill_manual(values = c("0" = "bisque1",
                               "1" = "brown4")) +
  theme(legend.position = c(0.7,0.8)) +
  facet_grid(~SL)

# Drift 10 - all TOL bands, facet grid
pivot2 <- tracks_SS_10 %>%
  pivot_longer(cols = starts_with("TOL")| starts_with("BB"), names_to = "SL", values_to = "Hz")    # Pivot longer for histogram facet grid

ggplot(data = pivot2, aes(Hz, fill = factor(BWpresence))) +
  geom_histogram(binwidth = 3, color = "black") +
  theme_classic() +
  ggtitle("Distribution of third octave level intensities: Drift 10") +
  scale_x_continuous(name = "Sound Intensity (dB)", breaks=seq(50,110,20)) +
  guides(fill = guide_legend(title = "Beaked Whale Presence")) +
  scale_fill_manual(values = c("0" = "bisque1",
                               "1" = "brown4")) +
  theme(legend.position = c(0.6,0.76)) +
  facet_grid(~SL)

# Drift 8 - TOL bands 125 and 2000 only
pivot3 <- tracks_SS_08 %>%
  dplyr::select(-TOL_63, -TOL_20000, -TOL_5000) %>%
  pivot_longer(cols = starts_with("TOL"), names_to = "TOL", values_to = "Hz")    # Pivot longer for histogram facet grid

TOL.labs <- c(`TOL_125` = "TOL at 125 Hz", 
              `TOL_2000` = "TOL at 2000 Hz")
```

```{r drift-8-histogram}
ggplot(data = pivot3, aes(Hz, fill = factor(BWpresence))) +
  geom_histogram(binwidth = 1, color = "black") +
  theme_classic() +
  ggtitle("Distribution of Third-Octave Level (TOL) Intensities: Buoy 8") +
  scale_x_continuous(name = "Sound Intensity (dB)") +
  scale_fill_manual(name = "Beaked Whale Presence", 
                    labels = c("0" = "Absent", "1" = "Present"),
                    values = c("0" = "bisque1",
                               "1" = "brown4")) +
  theme(legend.position = c(0.6,0.7)) +
  facet_wrap(~TOL, labeller = as_labeller(TOL.labs))
```

```{r}
# Drift 10 - TOL bands 125 and 2000 only
pivot4 <- tracks_SS_10 %>%
  dplyr::select(-TOL_63, -TOL_20000, -TOL_5000) %>%
  pivot_longer(cols = starts_with("TOL"), names_to = "TOL", values_to = "Hz")    # Pivot longer for histogram facet grid

ggplot(data = pivot4, aes(Hz, fill = factor(BWpresence))) +
  geom_histogram(binwidth = 1, color = "black") +
  theme_classic() +
  ggtitle("Distribution of Third-Octave Level (TOL) Intensities: Buoy 10") +
  scale_x_continuous(name = "Sound Intensity (dB)") +
  scale_fill_manual(name = "Beaked Whale Presence", 
                    labels = c("0" = "Absent", "1" = "Present"),
                    values = c("0" = "bisque1",
                               "1" = "brown4")) +
  theme(legend.position = c(0.6,0.7)) +
  facet_wrap(~TOL, labeller = as_labeller(TOL.labs))

TOL.labs <- c(`TOL_125` = "TOL at 125 Hz", 
              `TOL_2000` = "TOL at 2000 Hz")
```

Scatterplots

```{r}
# SST vs TOL 2000
ggplot(data = tracks_SS_08, aes(x = sst_mean, y = `TOL_2000`)) +
  geom_point(aes(color = species)) + scale_color_discrete(na.translate=FALSE)

# SST over time
ggplot(tracks_SS_08, aes(x = UTC, y = sst_mean)) +
  geom_line() +
  geom_point(aes(color = species)) +
  scale_color_discrete(na.translate=FALSE)
  #geom_smooth(method = "loess")

# Depth vs TOL 125
ggplot(data = tracks_SS_08, aes(x = depth, y = `TOL_125`)) +
  geom_line() +
  geom_point(aes(color = species)) +
  scale_color_discrete(na.translate=FALSE)

#Mean depth over time
ggplot(means_08, aes(x = `as.Date(UTC)`, y = depth*(-1))) +
  geom_line()

# Chlorophyll vs TOL 125
ggplot(data = tracks_SS_08, aes(x = chlorophyll_mean, y = `TOL_125`)) +
  geom_point(aes(color = species))

# CHl over time
ggplot(tracks_SS_08, aes(x = UTC, y = chlorophyll_mean)) +
  geom_line() +
  geom_point(aes(color = species)) +
  scale_color_discrete(na.translate=FALSE)
  #geom_smooth(method = "loess")

# Mixed Layer Depth vs TOL125 with species
ggplot(data = tracks_SS_08, aes(x = mldDepth, y = `TOL_125`)) +
  geom_point(aes(color = species))

# Mixed Layer Depth over time
ggplot(tracks_SS_08, aes(x = UTC, y = mldDepth*(-1))) +
  geom_line() +
  geom_point(aes(color = species)) +
  scale_color_discrete(na.translate=FALSE)

ggplot(data = tracks_SS_08, aes(x = slope, y = `TOL_5000`)) +
  geom_point(aes(color = species))
```

Facet Grids for Env. Variables
```{r}
#For surface/depth variables USING DAILY MEAN VARIABLES
# See above for means_08 derivation
track08Long <- means_08 %>%
  pivot_longer(cols = sst_mean:ssh_mean,
               names_to="parameter",
               values_to="value") %>%
  filter(parameter %in% c("sst_mean", "dist_shore", "chlorophyll_mean", "ssh_mean")) %>%
  rename(Date = `as.Date(UTC)`)

p <- ggplot(data = track08Long, aes(x = Date, y = value)) +
  geom_line()
 # geom_point(aes(size = BWpresence))   ### How best to represent number of BW detections per day?

p + facet_grid(parameter ~ ., scales = "free_y")

# For MLD/Thermocline variables 
track08Long2 <- tracks_SS_08 %>%
  pivot_longer(cols = mldDepth:sal400,
               names_to="parameter",
               values_to="value") %>%
  filter(parameter %in% c("mldDepth", "mldTemp", "ttDepth", "ttTemp", "temp400", "sal400")) %>%
    rename(Date = UTC) %>%
    subset(species!="?BW")

q <- ggplot(data = track08Long2, aes(x = Date, y = value)) +
  geom_line() + geom_point(aes(color = species)) + scale_color_discrete(na.translate=FALSE)
q + facet_grid(parameter ~ ., scales = "free_y")

# For temp400 and salinity400 variables
env.labs <- c(`sal400` = "Salinity at 400m", 
              `temp400` = "Temperature at 400m")

track08Long3 <- tracks_SS_08 %>%
  pivot_longer(cols = temp400:sal400,
               names_to="parameter",
               values_to="value") %>%
  filter(parameter %in% c("temp400", "sal400")) %>%
  rename(`Date (2018)` = UTC)

q <- ggplot(data = track08Long3, aes(x = `Date (2018)`, y = value)) +
  geom_line() + 
  geom_point(aes(color = species)) + 
  scale_color_discrete(na.translate=FALSE, labels = c("ZC" = "Cuvier's", "BW43" = "BW43", "BW" = "BW"))
q + facet_grid(parameter ~ ., scales = "free_y", labeller = as_labeller(env.labs))
```

Co-occurrence plots of drift 8 and vessels 
```{r}
#### 09/04/18 co-occurrence of vessel and drift within 10 km
track08_09.04 <- tracks_SS_08 %>%
  filter(UTC >= "2018-09-04 00:00:00" & UTC <= "2018-09-05 00:00:00")

ggplot(data = track08_09.04, aes(x = UTC, y = `TOL_125`)) +
  geom_vline(xintercept = as.POSIXct("2018-09-04 14:13:00"), color = "yellow", linewidth = 2) +
  geom_text(aes(as.POSIXct("2018-09-04 14:13:00"), 88, label = "Enter vessel", hjust = 1.0)) +
  geom_vline(xintercept = as.POSIXct("2018-09-04 15:17:00"), color = "yellow", linewidth = 2) +
  geom_text(aes(as.POSIXct("2018-09-04 15:13:00"), 87, label = "Exit vessel", hjust = -0.1)) +
  geom_line() + 
  geom_point(data = track08_09.04, aes(color = species, size = 2)) +
  scale_color_discrete(na.translate=FALSE) +
  scale_size_continuous(guide = "none") +
  ggtitle("Sound levels during co-occurrence of Drift 8 and a vessel on 09-04-18\n(Within 7 km)") +
  scale_y_continuous(name = "Sound Intensity (dB) at TOL 125 Hz") +
  scale_color_discrete(na.translate=FALSE, labels = c("ZC" = "Cuvier's", "BW43" = "BW43"))

#### 09/05/18 co-occurrence of vessel and drift within 10 km.
track08_09.05 <- tracks_SS_08 %>%
  filter(UTC >= "2018-09-04 21:00:00" & UTC <= "2018-09-06 00:00:00")

ggplot(data = track08_09.05, aes(x = UTC, y = `TOL_125`)) +
  geom_vline(xintercept = as.POSIXct("2018-09-05 01:33:00"), color = "yellow", linewidth = 1) +
  geom_text(aes(as.POSIXct("2018-09-05 01:33:00"), 86, label = "Enter vessel", hjust = 1)) +
  geom_vline(xintercept = as.POSIXct("2018-09-05 02:08:00"), color = "yellow", linewidth = 1) +
  geom_text(aes(as.POSIXct("2018-09-05 02:08:00"), 85, label = "Exit vessel", hjust = -0.1)) +
  geom_line() + 
  geom_point(data = track08_09.05, aes(color = species, size = 2)) +
  scale_size_continuous(guide = "none") +
  ggtitle("Sound levels during co-occurrence of Drift 8 and a vessel on 09-05-18\n(Within 5 km)") +
  scale_y_continuous(name = "Sound Intensity (dB) at TOL 125 Hz") +
  scale_color_discrete(na.translate=FALSE, labels = c("ZC" = "Cuvier's", "BW43" = "BW43"))

#### 09/10/18 co-occurrence of vessel and drift within 10 km.
track08_09.10 <- tracks_SS_08 %>%
  filter(UTC >= "2018-09-10 00:00:00" & UTC <= "2018-09-11 00:00:00")

ggplot(data = track08_09.10, aes(x = UTC, y = `TOL_125`)) +
  geom_vline(xintercept = as.POSIXct("2018-09-10 12:58:00"), color = "yellow", linewidth = 1) +
  geom_text(aes(as.POSIXct("2018-09-10 12:58:00"), 86, label = "Enter vessel", hjust = 1)) +
  geom_vline(xintercept = as.POSIXct("2018-09-10 14:01:00"), color = "yellow", linewidth = 1) +
  geom_text(aes(as.POSIXct("2018-09-10 14:01:00"), 85, label = "Exit vessel", hjust = -0.1)) +
  geom_line() +
  geom_point(data = track08_09.10, aes(color = species, size = 2)) +
  scale_size_continuous(guide = "none") +
  ggtitle("Sound levels during co-occurrence of Drift 8 and a vessel on 09-10-18\n(Within 2 km)") +
  scale_y_continuous(name = "Sound Intensity (dB) at TOL 125 Hz") +
  scale_color_discrete(na.translate=FALSE, labels = c("ZC" = "Cuvier's", "BW43" = "BW43"))

#### 09/22/18 co-occurrence of vessel and drift within 10 km
track08_09.22 <- tracks_SS_08 %>%
  filter(UTC >= "2018-09-22 00:00:00" & UTC <= "2018-09-23 00:00:00")

ggplot(data = track08_09.22, aes(x = UTC, y = `TOL_125`)) +
  geom_vline(xintercept = as.POSIXct("2018-09-22 19:13:00"), color = "yellow", linewidth = 1) +
  geom_text(aes(as.POSIXct("2018-09-22 19:13:00"), 86, label = "Enter vessel", hjust = 1)) +
  geom_vline(xintercept = as.POSIXct("2018-09-22 19:34:00"), color = "yellow", linewidth = 1) +
  geom_text(aes(as.POSIXct("2018-09-22 19:34:00"), 85, label = "Exit vessel", hjust = -0.1)) +
  geom_line() + 
  geom_point(data = track08_09.22, aes(color = species, size = 2)) +
  scale_size_continuous(guide = "none") +
  ggtitle("Sound levels during co-occurrence of Drift 8 and a vessel on 09-22-18\n(Within 9 km)") +
  scale_y_continuous(name = "Sound Intensity (dB) at TOL 125 Hz") +
  scale_color_discrete(na.translate=FALSE, labels = c("ZC" = "Cuvier's", "BW43" = "BW43"))

#### 09/23/18 co-occurrence of vessel and drift within 10 km. NO whale detections in this time period.
track08_09.23 <- tracks_SS_08 %>%
  filter(UTC >= "2018-09-23 00:00:00" & UTC <= "2018-09-24 00:00:00")

ggplot(data = track08_09.23, aes(x = UTC, y = `TOL_125`)) +
  geom_vline(xintercept = as.POSIXct("2018-09-23 03:55:00"), color = "yellow", linewidth = 1) +
  geom_text(aes(as.POSIXct("2018-09-23 03:55:00"), 86, label = "Enter vessel", hjust = 1)) +
  geom_vline(xintercept = as.POSIXct("2018-09-23 04:23:00"), color = "yellow", linewidth = 1) +
  geom_text(aes(as.POSIXct("2018-09-23 04:23:00"), 85, label = "Exit vessel", hjust = -0.1)) +
  geom_line() +
  ggtitle("Sound levels during co-occurrence of Drift 8 and a vessel on 09-23-18\n(Within 6 km)") +
  scale_y_continuous(name = "Sound Intensity (dB) at TOL 125 Hz") +
  scale_color_discrete(na.translate=FALSE, labels = c("ZC" = "Cuvier's", "BW43" = "BW43"))
```

BW Occurrences
```{r}
# Filter Drift 8 by BW presence and absence
BW1 <- tracks_SS_08 %>%
  filter(BWpresence == 1)
BW0 <- tracks_SS_08 %>%
  filter(BWpresence == 0)

ggplot(data = tracks_SS_08, mapping = aes(TOL_125)) + geom_histogram(binwidth = 0.1, color = "black")

ggplot() + 
  geom_histogram(data = BW0, aes(TOL_5000), binwidth = 0.5, fill = "black", color = "white") +
  geom_histogram(data = BW1, aes(TOL_5000), binwidth = 0.5, fill = "yellow", color = "black")

# Percent of time BWs are detected
print(paste(nrow(BW1) / nrow(tracks_SS_08) * 100,"%"))

# Most abundant BW species when whale is detected (remove NA values)
tracks_SS_08 %>%
  filter(!is.na(species)) %>%
  ggplot() + geom_bar(aes(species))

# Sound levels BW are exposed to by species
ZC <- tracks_SS_08 %>%
  filter(species == "ZC")

BW43 <- tracks_SS_08 %>%
  filter(species == "BW43")

print(max(ZC$TOL_125))
print(max(BW43$TOL_125))

ggplot() + geom_histogram(data = ZC, mapping = aes(TOL_125), binwidth = 0.2, fill = "brown", color = "white") +
  geom_histogram(data = BW43, mapping = aes(TOL_125), binwidth = 0.2, fill = "yellow", color = "white") +
  theme_classic()
```

Density Plots - Species at TOL sound levels
```{r}
# Broadband
tracks_SS_08 %>%
  filter(!is.na(species)) %>%
  ggplot(aes(`BB_20.24000`, fill = species)) + 
  geom_density(alpha = 0.5) +
  ggtitle("Broadband") +
  scale_x_continuous(name = "Sound Intensity (dB)") +
  scale_fill_discrete(labels = c("ZC" = "Cuvier's", "BW43" = "BW43"))

# TOL 63
tracks_SS_08 %>%
  filter(!is.na(species)) %>%
  ggplot(aes(TOL_63, fill = species)) + 
  geom_density(alpha = 0.5) +
  ggtitle("63 Hz") +
  scale_x_continuous(name = "Sound Intensity (dB)") +
  scale_fill_discrete(labels = c("ZC" = "Cuvier's", "BW43" = "BW43"))

# TOL 125
tracks_SS_08 %>%
  filter(!is.na(species)) %>%
  ggplot(aes(TOL_125, fill = species)) + 
  geom_density(alpha = 0.5) +
  ggtitle("125 Hz") +
  scale_x_continuous(name = "Sound Intensity (dB)") +
  scale_fill_discrete(labels = c("ZC" = "Cuvier's", "BW43" = "BW43"))

# TOL 2000
tracks_SS_08 %>%
  filter(!is.na(species)) %>%
  ggplot(aes(TOL_2000, fill = species)) + 
  geom_density(alpha = 0.5) +
  ggtitle("2000 Hz") +
  scale_x_continuous(name = "Sound Intensity (dB)") +
  scale_fill_discrete(labels = c("ZC" = "Cuvier's", "BW43" = "BW43"))

# TOL 5000
tracks_SS_08 %>%
  filter(!is.na(species)) %>%
  ggplot(aes(TOL_5000, fill = species)) + 
  geom_density(alpha = 0.5) +
  ggtitle("5000 Hz") +
  scale_x_continuous(name = "Sound Intensity (dB)") +
  scale_fill_discrete(labels = c("ZC" = "Cuvier's", "BW43" = "BW43"))

# TOL 20,000
tracks_SS_08 %>%
  filter(!is.na(species)) %>%
  ggplot(aes(TOL_20000, fill = species)) + 
  geom_density(alpha = 0.5) +
  ggtitle("20,000 Hz") +
  scale_x_continuous(name = "Sound Intensity (dB)") +
  scale_fill_discrete(labels = c("ZC" = "Cuvier's", "BW43" = "BW43"))
```

Histograms with BW presence/absence and environmental variables
```{r}
# SST
ggplot(data = tracks_SS_08, aes(sst_mean, fill = factor(BWpresence))) +
  geom_histogram(binwidth = 0.1, color = "black")

# Depth
ggplot(data = tracks_SS_08, aes(depth, fill = factor(BWpresence))) +
  geom_histogram(binwidth = 10, color = "black")

# Bottom slope
ggplot(data = tracks_SS_08, aes(slope, fill = factor(BWpresence))) +
  geom_histogram(binwidth = 0.005, color = "black")

# SSH
ggplot(data = tracks_SS_08, aes(ssh_mean, fill = factor(BWpresence))) +
  geom_histogram(binwidth = 0.005, color = "black")

# Chlorophyll
ggplot(data = tracks_SS_08, aes(chlorophyll_mean, fill = factor(BWpresence))) +
  geom_histogram(binwidth = 0.005, color = "black")

# Mixed layer depth Depth
ggplot(data = tracks_SS_08, aes(mldDepth, fill = factor(BWpresence))) +
  geom_histogram(binwidth = 0.25, color = "black")

# Mixed layer depth Temperature
ggplot(data = tracks_SS_08, aes(mldTemp, fill = factor(BWpresence))) +
  geom_histogram(binwidth = 0.1, color = "black")

# Thermocline Depth
ggplot(data = tracks_SS_08, aes(ttDepth, fill = factor(BWpresence))) +
  geom_histogram(binwidth = 0.25, color = "black")

#Thermocline temperature
ggplot(data = tracks_SS_08, aes(ttTemp, fill = factor(BWpresence))) +
  geom_histogram(binwidth = 0.05, color = "black")

# Temperature at 400m
ggplot(data = tracks_SS_08, aes(temp400, fill = factor(BWpresence))) +
  geom_histogram(binwidth = 0.01, color = "black")

# Salinity at 400 m
ggplot(data = tracks_SS_08, aes(sal400, fill = factor(BWpresence))) +
  geom_histogram(binwidth = 0.001, color = "black")
```

### Plot maps for Powerpoint images

Set Parameters
```{r}
theme_set(theme_bw())
tmap_mode("plot")
```

Create sf objects and basemaps for maps (not needed for environmental variables)
```{r}
whales_8_10sf <- BWsf %>%                            
dplyr::filter(station == "8" | station == "10")

trackssf <- st_as_sf(tracks, coords = c("Longitude", "Latitude"), crs=4326)
trk_TOL_08 <- st_as_sf(tracks_SS_08, coords = c("Longitude", "Latitude"), crs=4326)
trk_TOL_10 <- st_as_sf(tracks_SS_10, coords = c("Longitude", "Latitude"), crs=4326)
trackBase <- get_tiles(whales_8_10sf, provider = "Esri.WorldShadedRelief")
allTrackbase <- get_tiles(trackssf, provider = "Stamen.TonerLite", zoom = 5)
```

Map of whole study area
```{r}
# Establish bounding box
bbox <- st_bbox(c(xmin = -110, 
            xmax = -128, 
            ymax = 44, 
            ymin = 30), crs = st_crs(4326))

#Establish basemap
allTrackbase <- get_tiles(bbox, provider = "Esri.WorldGrayCanvas", zoom = 5)

# Establish inset map
us_region <- st_bbox(c(xmin = -90,
            xmax = -120, 
            ymax = 48, 
            ymin = 25), crs = st_crs(4326))
usa <- get_tiles(us_region, provider = "Esri.WorldGrayCanvas", zoom = 4)

sg <- bb_poly(bb(tracks_sf))
inset <- tm_shape(usa) + tm_rgb() + tm_shape(sg) + tm_borders(col = "red")

>>>>

---
title: " extracting environmenal and oceanographic variables"
output: html_document
date: "2022-11-06"
---

@@

## This only downloads 11 of 13 successfully - FIX
for (t in tracklist) {
  Hdata <- matchEnvData(as.data.frame(get(paste0("tracks_SS_", substr(t, 11, 12)))), nc=PAMmisc::hycomList, var=c('water_temp', 'salinity'), raw=TRUE, depth = 0:1000)

  thermData <- bind_rows(lapply(Hdata, function(x) {
    # do thermo calcs
    result <- calcThermoVars(x$matchDepth, x$water_temp)
    # get specific values
    result$temp400 <- x$water_temp[x$matchDepth == 400]
    result$sal400 <- x$salinity[x$matchDepth == 400]
    result
}))
  bindr <- cbind(get(paste0("tracks_SS_", substr(t, 11, 12))), thermData)
  assign(paste0("tracks_SS_", substr(t, 11, 12)), bindr)
}
```

```{r}
## HYCOM function above stops at buoys 07 and 08, need to manually run it here. FIX!!
test08 <- matchEnvData(as.data.frame(tracks_SS_08), nc=PAMmisc::hycomList, var=c('water_temp', 'salinity'), raw=TRUE, depth = 0:1000)

@@

Save final buoy df for analysis use next.
```{r}
saveRDS(tracks_SS_10, 'data/fulltrack10.rda')
saveRDS(tracks_SS_12, 'data/fulltrack12.rda')
saveRDS(tracks_SS_13, 'data/fulltrack13.rda')
saveRDS(tracks_SS_14, 'data/fulltrack14.rda')
saveRDS(tracks_SS_16, 'data/fulltrack16.rda')
saveRDS(tracks_SS_18, 'data/fulltrack18.rda')
saveRDS(tracks_SS_19, 'data/fulltrack19.rda')
saveRDS(tracks_SS_20, 'data/fulltrack20.rda')
saveRDS(tracks_SS_21, 'data/fulltrack21.rda')
saveRDS(tracks_SS_22, 'data/fulltrack22.rda')
saveRDS(tracks_SS_23, 'data/fulltrack23.rda')
saveRDS(tracks_SS_07, 'data/fulltrack07.rda')
saveRDS(tracks_SS_08, 'data/fulltrack08.rda')
```
